[Result, but poetic]: Does LLM Use Improve Data Science Education?

### 2024-08-02 paper/paper.qmd
worked more on NLP

---
title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
author: 
  - Rohan Alexander
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
date: today
date-format: long
abstract: "An abstract should eventually go here!!"
format: pdf
number-sections: true
bibliography: references.bib
nocite: | #references.bib entries at end, regardless of intext citation
  @* 
---

@@

library(tidyverse)
library(knitr)
library(kableExtra)
library(lintr)
library(tidyr)
library(tidytext)
library(textdata)
library(wordcloud)
library(reshape2)

@@

The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines. Their quick adoption in academic settings has created excitement and apprehension among educators and students alike. LLMs offer unprecedented support for students in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes. At the same time, their use raises important questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.

@@

All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat]. This led to a final dataset containing 121 responses across 41 variables that corresponded to each survey question.

@@

plot_data |>
  ggplot(aes(x = mins_complete)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Survey Response Times",
    x = "Time (Mins)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5)
  )
```
@fig-response-time-hist shows the distribution of response times to the end-of-course survey. Respondents who consented generally took between 0-40 mins to answer all of the questions, with most response times ranging between 2-10 minutes. There was one consenting respondent who took 4308 minutes, whose observation was dropped from the visualization. This suggests that students generally took a meaningful amount of time to engage with the survey.

```{r fig-gpa-hist}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Distribution of Student GPA
#| fig-align: center

@@

survey_data |>
  ggplot(aes(x = gpa)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Student GPA",
    x = "GPA",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.6)
  )
```

@@

Get inspired by Cahill and McCabe's paper for visualizations etc.

```{r tbl-year-dist}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Distribution of Students by Year
#| fig-align: center

@@

ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(
    title = "Student Distribution by Program of Study",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  scale_fill_discrete(name = "Program") +
  theme(
    axis.text.x = element_blank(), # Remove x-axis text
    axis.title.x = element_blank(), # Remove x-axis title
    plot.title = element_text(hjust = 1.6)
  ) +
  geom_text(aes(label = n),
    position = position_stack(vjust = 0.5)
  )
```

@@

table_data |>
  ggplot(aes(x = "", y = n, fill = ai_familiarity)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(
    title = "Student Familiarity with AI",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  scale_fill_discrete(name = "Program") +
  theme(
    axis.text.x = element_blank(), # Remove x-axis text
    axis.title.x = element_blank(), # Remove x-axis title
    plot.title = element_text(hjust = 1.35)
  ) +
  geom_text(aes(label = n),
    position = position_stack(vjust = 0.5)
  )
```

Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs. This overrepresentation of familiarity also primarily driven by the fact that the course is primarily for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, which could likely lead to a skewing of interest/affinity for computers.

@@

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me\"",
      Statements == "i_like_to_write" ~ "\"I like to write\"",
      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\n to be a good writer\"",
      TRUE ~ Statements
    ),
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(
    title = "Student Self-Perceptions on Writing",
    x = "Response",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
  )

multi_bar

# TODO: Fix label, Fix Legend title, Fix legibility of Response
```

@@

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\n catch my own mistakes\"",
      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\n sharing my writing\"",
      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\n my own writing ability\"",
      TRUE ~ Statements
    ),
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(
    title = "Student Self-Perceptions on Writing",
    x = "Response",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
  )

@@

@@

```{r fig-ai-use-multibar}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Hello
#| fig-align: center

# Create a multi-bar plot for AI use/perceptions grouped by response

# THROWS ERROR: FIX

# Select first three statements
writing_perceptions <- survey_data |>
  select(
    used_ai_any_reason,
    used_ai_in_sta302
  )

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
      TRUE ~ Statements
    ),
  ) |>
  filter(!is.na(Response)) |>
  filter(Response == "No" | Response == "Yes")

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(
    title = "Student Self-Perceptions of Coding",
    x = "Response",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
  )

@@

# Create a bar plot of how students used ChatGPT - FIX: sort from largest to smallest

type_data <- survey_data |>
  select(
    not_used,
    used_technical_q,
    used_conversation,
    used_general_knowledge,
    used_solving_hmk,
    used_check_soln,
    used_quick_q,
    used_expl_concepts,
    used_writing_content,
    used_writing_code
  ) |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(Statements = case_when(
    Statements == "not_used" ~ "Not Used",
    Statements == "used_technical_q" ~ "Technical Questions",
    Statements == "used_conversation" ~ "Fun Conversation",
    Statements == "used_general_knowledge" ~ "General Knowledge",
    Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
    Statements == "used_check_soln" ~ "Checking my Solutions",
    Statements == "used_quick_q" ~ "Quick Questions",
    Statements == "used_expl_concepts" ~ "Explaining Concepts",
    Statements == "used_writing_content" ~ "Writing Paper Content",
    Statements == "used_writing_code" ~ "Writing Paper Code",
    TRUE ~ Statements
  )) |>
  filter(Response == 1) |>
  count(Statements) |>
  arrange(desc(n))

type_data |>
  ggplot(aes(x = n, y = reorder(Statements, n))) +
  geom_bar(stat = "identity", fill = "grey50") +
  labs(title = "How Students Used ChatGPT", x = "Count", y = "Use Type") +
  theme_classic()
```

To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey. @fig-AI-usage-type-comp demonstrates that asking technical questions and explaining concepts were the two top use cases among students. Less than half used it to entirely solve their homework, which could be explained by the complex nature of the homework, that involved doing extensive research and writing. Also just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.

@@

#get positive words from NRC lexicon
neg_words <- get_sentiments("bing") |>
  filter(sentiment == "negative")

#get negative words from NRC lexicon
pos_words <- get_sentiments("bing") |>
  filter(sentiment == "positive")
```

We then performed sentiment analysis on the free-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.

@@

In the survey, students were asked to respond to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".

@@

# put comments into tidytext format
tidy_ai_comments <- ai_comments |>
  group_by(id) |>
  unnest_tokens(word, comment)

@@

```{r tbl-ai-appropriate}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Comparing Incidences of Positive and Negative Words.
#| tbl-subcap: ["Top 10 Mentioned Positive Words", "Top 10 Mentioned Negative Words"]
#| layout-ncol: 2
#| fig-align: center

@@

# Generate comparative word cloud
comparison.cloud(
  comparison_data,
  colors = c("darkred", "darkgreen"),
  max.words = 100,
  title.size = 1.5,
  scale = c(3, 0.5)  # Adjust scale for word size
)

@@

### Recommendations in STA302 
```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Hello
#| fig-align: center

recs_comments <- survey_data |>
  select(
    id,
    recs_for_ai_sta302
  )
```

>>>>

---
title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
author: 
  - Rohan Alexander
  - Luca Carnegie
thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
date: today
date-format: long
abstract: "An abstract should eventually go here!!"
format: pdf
number-sections: true
table-of-contents: true
bibliography: references.bib
nocite: | #references.bib entries at end, regardless of intext citation
  @* 
---

@@

library(tidyverse)
library(knitr)
library(kableExtra)
library(lintr)
library(tidyr)
library(tidytext)
library(textdata)
library(quanteda)
library(wordcloud)
library(reshape2)
library(topicmodels)

@@

The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines, creating excitement and apprehension among educators and students alike. They offer unprecedented support for student learning in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes. At the same time, their use raises important questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.

@@

All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat]. This led to a final dataset containing 121 unique responses to 18  survey questions.

@@

plot_data |>
  ggplot(aes(x = mins_complete)) +
  geom_histogram(binwidth = 2, fill = "lightblue", color = "black") +
  labs(
    title = "Distribution of Survey Response Times",
    x = "Time (Mins)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5)
  )
```
@fig-response-time-hist shows the distribution of response times to the end-of-course survey. Respondents who consented generally took between 0-40 mins to answer all of the questions, with most response times ranging between 2-10 minutes. There was one consenting respondent who took 4308 minutes, whose observation was dropped from the visualization. Overall, however, this suggests that students generally took a meaningful amount of time and thought to engage with the survey's questions.

```{r fig-gpa-hist}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Distribution of Student GPA
#| fig-align: center

@@

survey_data |>
  ggplot(aes(x = gpa)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(
    title = "Distribution of Student GPA",
    x = "GPA",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.6)
  )
```

@@

```{r tbl-year-dist}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Count of Students by Year
#| fig-align: center

@@

ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(
    title = "Distribution of Students by Program of Study",
    subtitle = "Visualizing the proportion of students across different programs",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  scale_fill_discrete(name = "Program of Study") +
  theme(
    axis.text.x = element_blank(), # Remove x-axis text
    axis.title.x = element_blank(), # Remove x-axis title
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  geom_text(aes(label = n),
    position = position_stack(vjust = 0.5)
  )
```

@@

# Custom colors
familiarity_colors <- c(
  "Not familiar" = "tomato",
  "Somewhat familiar" = "lightblue",
  "Very familiar" = "lightgreen"
)

table_data |>
  ggplot(aes(x = "", y = n, fill = ai_familiarity)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(
    title = "Student Familiarity with AI",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  scale_fill_manual(name = "Familiarity Level", values = familiarity_colors) +
  theme(
    axis.text.x = element_blank(), # Remove x-axis text
    axis.title.x = element_blank(), # Remove x-axis title
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  geom_text(aes(label = n),
    position = position_stack(vjust = 0.5)
  )
```

Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs. This overrepresentation of familiarity also primarily driven by the fact that the course is primarily targeted for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, possibly leading to a sample that is skewed to being more interested in computing.

@@

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me\"",
      Statements == "i_like_to_write" ~ "\"I like to write\"",
      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer\"",
      TRUE ~ Statements
    ),
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

# Custom colors
response_colors <- c(
  "A lot like me" = "lightgreen",
  "Somewhat describes me" = "lightblue",
  "Very different from me" = "tomato"
)

# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
  myPlot +
    guides(
      shape = guide_legend(override.aes = list(size = pointSize)),
      color = guide_legend(override.aes = list(size = pointSize)),
      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
    ) +
    theme(
      legend.title = element_text(size = textSize),
      legend.text = element_text(size = textSize),
      legend.key.size = unit(spaceLegend, "lines")
    )
}

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Response)) +
  geom_bar(position = "dodge", stat = "count") +
  facet_wrap(~Statements) +
  labs(
    title = "Student Self-Perceptions on Writing",
    subtitle = "Responses to statements about writing ease, enjoyment, and importance",
    x = "Response",
    y = "Number of Students"
  ) +
  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)

@@

writing_perceptions_long <- writing_perceptions |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(
    Statements = case_when(
      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\ncatch my own mistakes\"",
      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing\"",
      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\nmy own writing ability\"",
      TRUE ~ Statements
    ),
    Response = case_when(
      Response == "This is a lot like me" ~ "A lot like me",
      Response == "This is very different to me" ~ "Very different from me",
      Response == "This somewhat describes me" ~ "Somewhat describes me",
    )
  ) |>
  filter(!is.na(Response))

# Custom colors
response_colors <- c(
  "A lot like me" = "lightgreen",
  "Somewhat describes me" = "lightblue",
  "Very different from me" = "tomato"
)

# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
  myPlot +
    guides(
      shape = guide_legend(override.aes = list(size = pointSize)),
      color = guide_legend(override.aes = list(size = pointSize)),
      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
    ) +
    theme(
      legend.title = element_text(size = textSize),
      legend.text = element_text(size = textSize),
      legend.key.size = unit(spaceLegend, "lines")
    )
}

# Make the plot
multi_bar2 <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Response)) +
  geom_bar(position = "dodge", stat = "count") +
  facet_wrap(~Statements) +
  labs(
    title = "Student Self-Perceptions on Writing",
    subtitle = "Responses to statements about confidence and editing in writing",
    x = "Response",
    y = "Number of Students"
  ) +
  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_blank(), # Remove x-axis labels
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

multi_bar2 <- addSmallLegend(multi_bar2, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)

multi_bar2

@@

# Custom colors
response_colors <- c(
  "A lot like me" = "lightgreen",
  "Somewhat describes me" = "lightblue",
  "Very different from me" = "tomato"
)

# Make the plot
multi_bar <- writing_perceptions_long |>
  ggplot(aes(x = Response, fill = Statements)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(
    title = "Student Self-Perceptions of Coding",
    x = "Response",
    y = "Number of Students"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    axis.text.x = element_blank(),
  )

@@

@@

type_data <- survey_data |>
  select(
    not_used,
    used_technical_q,
    used_conversation,
    used_general_knowledge,
    used_solving_hmk,
    used_check_soln,
    used_quick_q,
    used_expl_concepts,
    used_writing_content,
    used_writing_code
  ) |>
  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
  mutate(Statements = case_when(
    Statements == "not_used" ~ "Not Used",
    Statements == "used_technical_q" ~ "Technical Questions",
    Statements == "used_conversation" ~ "Fun Conversation",
    Statements == "used_general_knowledge" ~ "General Knowledge",
    Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
    Statements == "used_check_soln" ~ "Checking Solutions",
    Statements == "used_quick_q" ~ "Quick Questions",
    Statements == "used_expl_concepts" ~ "Explaining Concepts",
    Statements == "used_writing_content" ~ "Writing Paper Content",
    Statements == "used_writing_code" ~ "Writing Paper Code",
    TRUE ~ Statements
  )) |>
  filter(Response == 1) |>
  count(Statements) |>
  arrange(desc(n))

type_data |>
  ggplot(aes(x = n, y = reorder(Statements, n))) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "How Students Used ChatGPT", x = "Number of Students", y = "Type of Usage") +
  theme_classic()
```

To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey. As we can see in @fig-AI-usage-type-comp, technical questions and explaining concepts were the two top use cases among students. More than half of students also used ChatGPT for Quick Questions, General Knowledge, Writing Paper Code, and Checking Solutions. Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.

@@

#get positive words from tidytext
neg_words <- get_sentiments() |>
  filter(sentiment == "negative")

#get negative words from tidytext
pos_words <- get_sentiments() |>
  filter(sentiment == "positive")
```

We then performed basic sentiment analysis on the open-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.

@@

In the survey, students were asked to provide an open-ended response to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?". These responses were attempted to be understood thematically.

@@

# put comments into tidytext format
tidy_ai_comments <- ai_comments |>
  group_by(id) |>
  unnest_tokens(word, comment) |>
  mutate(word = str_to_lower(string = word))

@@

# Generate comparative word cloud
comparison.cloud(
  comparison_data,
  colors = c("darkred", "darkgreen"),
  max.words = 100,
  title.size = 1.5,
  scale = c(3, 0.5)  # Adjust scale for word size
)
```

```{r tbl-ai-appropriate}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Comparing Incidences of Positive and Negative Words.
#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
#| layout-ncol: 2
#| fig-align: center

@@

### AI Implementation Recommendations

```{r recs-setup}
#| message: false
#| echo: false
#| warning: false

# Example sentiments data frame
sentiments <- get_sentiments()

# Prepare data
recs_comments <- survey_data |>
  select(id, recs_for_ai_sta302) |>
  mutate(comment = recs_for_ai_sta302) |>
  select(-c(recs_for_ai_sta302))

# Tokenize and clean individual words
tidy_recs_comments <- recs_comments |>
  ungroup() |>
  unnest_tokens(word, comment) |>
  anti_join(stop_words)

# Count word frequencies
word_counts <- tidy_recs_comments |>
  count(word, sort = TRUE)

# Bigram analysis
bigrams <- recs_comments |>
  ungroup() |>
  unnest_tokens(bigram, comment, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word & !word2 %in% stop_words$word) |>
  count(word1, word2, sort = TRUE) |>
  unite(bigram, word1, word2, sep = " ")

# Prepare data for comparison cloud (including both words and bigrams)
comparison_data_words <- tidy_recs_comments |>
  inner_join(sentiments, by = "word") |>
  count(word, sentiment, sort = TRUE)

comparison_data_bigrams <- bigrams |>
  separate(bigram, into = c("word1", "word2"), sep = " ") |>
  inner_join(sentiments, by = c("word1" = "word")) |>
  inner_join(sentiments, by = c("word2" = "word"), suffix = c(".word1", ".word2")) |>
  unite(bigram, word1, word2, sep = " ") |>
  count(bigram, sentiment.word1, sort = TRUE) |>
  rename(sentiment = sentiment.word1)

# Combine both data sets
combined_data <- bind_rows(
  comparison_data_words |> rename(term = word),
  comparison_data_bigrams |> rename(term = bigram)
) |>
  arrange(desc(n))

@@

```{r recs-wordcloud}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Word Cloud for Recommendations Comments
#| fig-align: center

combined_data |>
  with(
    wordcloud(
      words = term,
      freq = n,
      scale = c(3, 0.5),
      min.freq = 1, # Adjust scale as needed
      max.words = 100,    # Maximum number of words to display
      colors = "blue"     # Color for the word cloud
    )
  )

```{r tbl-recs-relpolarity}
#| message: false
#| echo: false
#| warning: false
#| tbl-cap: Comparing Incidences of Positive and Negative Words in Student.
#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
#| layout-ncol: 2
#| fig-align: center

# Generate positive words table
tbl_recs_pos <- tidy_recs_comments |>
  ungroup() |>
  select(-c("id")) |>
  inner_join(pos_words) |>
  count(word, sort = TRUE) |>
  head(10) |>
  kable(, format = "markdown")

# Generate negative words table
tbl_recs_neg <- tidy_recs_comments |>
  ungroup() |>
  select(-c("id")) |>
  inner_join(neg_words) |>
  count(word, sort = TRUE) |>
  head(10) |>
  kable(, format = "markdown")

tbl_recs_pos
tbl_recs_neg
```

```{r word-cloud}
#| message: false
#| echo: false
#| warning: false

# Word cloud based on bigrams
word_counts |>
  with(
    wordcloud(
      words = word,
      freq = n,
      scale = c(3, 0.5),  # Adjust scale as needed
      max.words = 100,    # Maximum number of words to display
      colors = "blue"     # Color for the word cloud
    )
  )
```
